
---
title: "Gear Ratios in MachCalc"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Gear Ratios in MachCalc}
  %\VignetteEngine{knitr::rmarkdown}
  \usepackage[utf8]{inputenc}
---

<!--
% \VignetteEngine{knitr::knitr}
% \VignetteIndexEntry{Introduction to  MachCalc}
-->
This vignette presents an overview of the main 
functions implemented in the MachCalc package. 


# Overview

This package was writtent mainly for my own use, but 
I hope others will also find a use for them and let 
me know if they see errors.

# List of main R functions and objects


| Function                      | Summary Description                                                 |
|-------------------------------|---------------------------------------------------------------------|
| `bestGearRatio`               | Finds pairs of gears producing ratios close to the one required     |
| `bore3P`                      | Size of the bore of a hole in which 3 given pins fit cotangentially |
| `gearRatio`                   | Support function for *bestGearRatio*
| `linesCircle`                 | Support function for *bore3P*
| `plotCircle`                  | Support function for *bore3P*
| `spindleSpeed`                | Calculates spindle speed


To begin, we'll load **machcalc** and call
the *bestGearRatio* function.

```{r setup, message = FALSE}
library(MachCalc)
bestGearRatio(1.27,100)
```

## `arrange()`: Arrange rows by variables

`dplyr::arrange()` orders the rows of a data frame by the values of one or more columns: 


The `desc()` helper allows you to order selected variables in descending order:


We can replicate in base R by using `[` with `order()`:

```{r}
mtcars[order(mtcars$cyl, mtcars$disp), , drop = FALSE]
```

Note the use of `drop = FALSE`. If you forget this, and the input is a data frame with a single column, the output will be a vector, not a data frame. This is a source of subtle bugs.

Base R does not provide a convenient and general way to sort individual variables in descending order, so you have two options:

* For numeric variables, you can use `-x`.
* You can request `order()` to sort all variables in descending order.

```{r, results = FALSE}
mtcars[order(mtcars$cyl, mtcars$disp, decreasing = TRUE), , drop = FALSE]
mtcars[order(-mtcars$cyl, -mtcars$disp), , drop = FALSE]
```

## `distinct()`: Select distinct/unique rows

`dplyr::distinct()` selects unique rows:


There are two equivalents in base R, depending on whether you want the whole data frame, or just selected variables:

## `filter()`: Return rows with matching conditions

`dplyr::filter()` selects rows where an expression is `TRUE`:


The closest base equivalent (and the inspiration for `filter()`) is `subset()`:


You can also use `[` but this also requires the use of `which()` to remove `NA`s:


## `mutate()`: Create or transform variables

`dplyr::mutate()` creates new variables from existing variables:


The closest base equivalent is `transform()`, but note that it cannot use freshly created variables:

Alternatively, you can use `$<-`:

```{r}
mtcars$cyl2 <- mtcars$cyl * 2
mtcars$cyl4 <- mtcars$cyl2 * 2
```

When applied to a grouped data frame, `dplyr::mutate()` computes new variable once per group:


To replicate this in base R, you can use `ave()`:


## `pull()`: Pull out a single variable

`dplyr::pull()` extracts a variable either by name or position:


This equivalent to `[[` for positions and `$` for names:

```{r}
mtcars[[1]]
mtcars$cyl
```


## `rename()`: Rename variables by name

`dplyr::rename()` allows you to rename variables by name or position:


Renaming variables by position is straight forward in base R:

```{r}
iris2 <- iris
names(iris2)[2] <- "sepal_width"
```

Renaming variables by name requires a bit more work:

```{r}
names(iris2)[names(iris2) == "Sepal.Length"] <- "sepal_length"
```

Subsetting variables by position is straightforward in base R:

```{r}
iris[1:3] # single argument selects columns; never drops
iris[1:3, , drop = FALSE]
```

You have two options to subset by name:

```{r}
iris[c("Species", "Sepal.Length")]
subset(iris, select = c(Species, Sepal.Length))
```

Subsetting by function of name requires a bit of work with `grep()`:

```{r}
iris[grep("^Petal", names(iris))]
```

And you can use `Filter()` to subset by type:

```{r}
Filter(is.factor, iris)
```

## `summarise()`: Reduce multiple values down to a single value

`dplyr::summarise()` computes one or more summaries for each group:


I think the closest base R equivalent uses `by()`. Unfortunately `by()` returns a list of data frames, but you can combine them back together again with `do.call()` and `rbind()`:

```{r}
mtcars_by <- by(mtcars, mtcars$cyl, function(df) {
  with(df, data.frame(cyl = cyl[[1]], mean = mean(disp), n = nrow(df)))
})
do.call(rbind, mtcars_by)
```

`aggregate()` comes very close to providing an elegant answer:

```{r}
agg <- aggregate(disp ~ cyl, mtcars, function(x) c(mean = mean(x), n = length(x)))
agg
```

But unfortunately while it looks like there are `disp.mean` and `disp.n` columns, it's actually a single matrix column:

```{r}
str(agg)
```

You can see a variety of other options at <https://gist.github.com/hadley/c430501804349d382ce90754936ab8ec>, many generated by the R community inspired by [a tweet](https://twitter.com/hadleywickham/status/1231252596712771585).

## `slice()`: Choose rows by position

`slice()` selects rows with their location:


This is straightforward to replicate with `[`:

```{r}
mtcars[25:nrow(mtcars), , drop = FALSE]
```

# Two-table verbs

When we want to merge two data frames, `x` and `y`), we have a variety of different ways to bring them together. Various base R `merge()` calls are replaced by a variety of dplyr `join()` functions.

| dplyr                  | base                                    |
|------------------------|-----------------------------------------|
| `inner_join(df1, df2)` |`merge(df1, df2)`                        | 
| `left_join(df1, df2) ` |`merge(df1, df2, all.x = TRUE)`          | 
| `right_join(df1, df2)` |`merge(df1, df2, all.y = TRUE)`          | 
| `full_join(df1, df2)`  |`merge(df1, df2, all = TRUE)`            | 
| `semi_join(df1, df2)`  |`df1[df1xx, , drop = FALSE]`  | 
| `anti_join(df1, df2)`  |`df1[!df1xx, , drop = FALSE]` | 

For more information about two-table verbs, see `vignette("two-table")`.

### Mutating joins

dplyr's `inner_join()`, `left_join()`, `right_join()`, and `full_join()` add new columns from `y` to `x`, matching rows based on a set of "keys", and differ only in how missing matches are handled. They are equivalent to calls to `merge()` with various settings of the `all`, `all.x`, and `all.y` arguments. The main difference is the order of the rows:

* dplyr preserves the order of the `x` data frame.
* `merge()` sorts the key columns.

### Filtering joins

dplyr's `semi_join()` and `anti_join()` affect only the rows, not the columns:


They can be replicated in base R with `[` and `%in%`:


Semi and anti joins with multiple key variables are considerably more challenging to implement.